---
title: "Tankerkoenig"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      fig.width = 9, fig.height = 7)

library(tidyverse)
library(sf)
library(tmap)
library(data.table)
library(plotly)
library(ggplot2)

```

### Prepare Data

The data was taken from Tankerkoenig for the month of July. We identified the top 5 brands in Germany and filtered for all of the stations under these brands. Then we grouped by station id and calculated the average price for diesel, e5 and e10. Since there were so many stations, we took a sample of the most expensive stations and plotted


```{r, cache=TRUE}
# Import tankstation data and make brands toupper()
stations <- fread("data/stations/stations.csv")
stations[, brand := toupper(brand)]

# Filter out bad coords and add 2-digit zip code
stations <- stations[longitude > 0 & latitude > 0]

# Extract Top 5 Brands
top_5 <- stations[, .(count = .N), by=brand][order(-count)][1:5, brand]
top_5_brands <- stations[brand %in% top_5]

# Import price data
prices <- fread("data/2019_Juli_Sommerferien.csv")
prices[, date := as.Date(date)]
station_price <- prices[, .(diesel = round(mean(diesel), 2),
                          e5     = round(mean(e5), 2),
                          e10    = round(mean(e10), 2)), 
                      by = .(station_uuid)]

# Join stations with prices
DT <- merge(top_5_brands, station_price, by.x="uuid", by.y="station_uuid", all.x=TRUE)

# Take only expensive stations
top_100_diesel <- DT[order(-diesel), uuid][1:100]
top_100_e5     <- DT[order(-e5), uuid][1:100]
top_100_e10    <- DT[order(-e10), uuid][1:100]

# Remove stations without price
DT <- DT[!is.na(date) & diesel > 1.0 & e5 > 1.0 & e10 > 1.0]

# Extracting longitude and latitude and transforming
station_map <- DT %>% 
    sf::st_as_sf(coords = c("longitude", "latitude"), crs=4326) %>% 
    sf::st_transform(crs=4326)
```


<br>

## Teuersten Tankstellen in Deutschland {.tabset}

```{r}
# tm map settings
tmap_mode("view")
popup <- c("brand", "street", "city", "post_code", "diesel", "e5", "e10")
```

### Diesel

#### Durschnittlicher Dieselpreis der Top 5 Marken im Juli 2019:`r round(mean(station_map$diesel), 2)` €

```{r}
station_map %>% 
    dplyr::filter(uuid %in% top_100_diesel) %>% 
    tm_shape() + 
    tm_dots(size = .07, col="diesel", style="cont", popup.vars=popup) +
    tm_view(basemaps = "Esri.WorldTopoMap")
```

### e5

#### Durschnittlicher e5-Preis der Top 5 Marken im Juli 2019:`r round(mean(station_map$e5), 2)` €

```{r}
station_map %>%
    dplyr::filter(uuid %in% top_100_e5) %>% 
    tm_shape() + 
    tm_dots(size = .07, col="e5", style="cont", popup.vars=popup) +
    tm_view(basemaps = "Esri.WorldTopoMap")
```

### e10

#### Durschnittlicher e10-Preis der Top 5 Marken im Juli 2019:`r round(mean(station_map$e10), 2)` €

```{r}
station_map %>% 
    dplyr::filter(uuid %in% top_100_e10) %>% 
    tm_shape() + 
    tm_dots(size = .07, col="e10", style="cont", popup.vars=popup) +
    tm_view(basemaps = "Esri.WorldTopoMap")
```

### Data

```{r}
top_100 <- DT[uuid %in% c(top_100_diesel, top_100_e5, top_100_e10),
              .(uuid, brand, city, diesel, e5, e10)]
DT::datatable(top_100)
```


<br>

## Top 5 for each category {.tabset}

```{r}
daily_price <- prices[, .(diesel = round(mean(diesel), 2),
                          e5     = round(mean(e5), 2),
                          e10    = round(mean(e10), 2)), 
                      by = .(date, station_uuid)]
# Join stations with prices
daily_price <- merge(top_5_brands, daily_price, by.x="uuid", by.y="station_uuid", all.x=TRUE)

daily_price <- daily_price[, .(diesel = round(mean(diesel), 2),
                          e5     = round(mean(e5), 2),
                          e10    = round(mean(e10), 2)), 
                      by = .(date, brand)]

daily_price[, date := as.Date(date)]
```


### Diesel

```{r}
# Average diesel overall
avg_diesel <- prices[, .(diesel = mean(diesel, na.rm = TRUE)), by=date]
avg_diesel[, date := as.Date(date)]

# Plot the price change for each station for the month of July
plt <- ggplot(data=NULL, aes(date, diesel)) +
    geom_line(data=avg_diesel, colour = "darkgrey", size = 3, alpha=.5) +
    geom_line(data=daily_price, aes(colour = brand)) +
    theme_bw()

ggplotly(plt)
```

### e5

```{r}
# Average diesel overall
avg_e5 <- prices[, .(e5 = mean(e5, na.rm = TRUE)), by=date]
avg_e5[, date := as.Date(date)]

# Plot the price change for each station for the month of July
plt <- ggplot(data=NULL, aes(date, e5)) +
    geom_line(data=avg_e5, colour = "darkgrey", size = 3, alpha=.5) +
    geom_line(data=daily_price, aes(colour = brand)) +
    theme_bw()

ggplotly(plt)
```

### e10

```{r}
# Average diesel overall
avg_e10 <- prices[, .(e10 = mean(e10, na.rm = TRUE)), by=date]
avg_e10[, date := as.Date(date)]

# Plot the price change for each station for the month of July
plt <- ggplot(data=NULL, aes(date, e10)) +
    geom_line(data=avg_e10, colour = "darkgrey", size = 3, alpha=.5) +
    geom_line(data=daily_price, aes(colour = brand)) +
    theme_bw()

ggplotly(plt)
```


<br>

## Where are these brands located

```{r, cache=10}
plz_gebiete <- read_sf("data/plz-2stellig/plz-2stellig.shp")

stations[, post_2stellig := substring(post_code, 1, 2)]

brand_zip <- stations %>% 
  filter(brand %in% top_5) %>%
  group_by(post_2stellig, brand) %>% 
  summarise(count = n()) %>% 
  ungroup()

zip_plt <- brand_zip %>% 
  left_join(plz_gebiete, by=c("post_2stellig" = "plz")) %>% 
  ggplot() +
  geom_sf(aes(geometry=geometry, fill=count)) +
  coord_sf(crs = st_crs(4326)) +
  facet_wrap(~ brand) +
  scale_fill_gradient2(low="white", mid="yellow", high="red", midpoint = 27) +
  hrbrthemes::theme_ipsum_tw()

zip_plt
```


